Implications of afforestation policies for global markets

### 2024-01-24 paper.Rmd
Description of combining the model

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland. Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

@@

South America and Sub-Saharan Africa are two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023]. While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. There are other nations that have a big potential for gains from afforestation, but they also face challenges. India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wa    stelands that may not be well suited for afforestation [@amjath2019]. Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.

@@

>>>>

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland. Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

@@

South America and Sub-Saharan Africa are two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023]. While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. There are other nations that have a big potential for gains from afforestation, but they also face challenges. India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019]. Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.

@@

# Connecting the GTAP and GTM models

The standard GTAP model represents a leading CGE model of the global economy with a specific focus on trade linkages. Its general equilibrium nature means that all markets that are considered in the model are cleared with market-clearing prices. The extensive coverage of sectors---the latest version of the GTAP database includes 68 sectors---makes it difficult to model sectors individually; instead the model applies nested CES production functions to all sectors, differentiating them by parameters only. To model one sector, e.g., forestry, differently from the rest would require major redesign of the model. While specific versions of the GTAP model are definitely possible but extremely common as they require to be be fully parametrized for all countries, which makes the option of adapting the GTAP model for a one-off application impractical.  To analyze carbon sequestration in GTAP, which would require a lot of effort, we instead create a connection between the standard GTAP model and the GTM model which has been specifically designed to analyze the relationship among land rents, forested land cover and carbon sequestration through forests, among other things.

Establishing a connection between the standard GTAP model and the GTM model is a two-step process. First, we decide on the set of common variables between the two models in those markets that are included in both models and their closure, i.e., which variables will be determined endogenously by one model and held as exogenous for the other model. In our case, we selected the return on land to be modeled by the GTAP model and fed into the GTM model as exogenous, while the amount of land available for crop and livestock production, while exogenous to the GTAP model, is endogenously determined by the GTM model. In the second, step we establish the connection between the models on the software level. This step requires that the GTAP model, written in GEMPACK can operate with the inputs from the GTM model, written in GAMS, and submit the outputs back to the GTM model. In order to allow for the two quite dated technologies, GEMPACK and GAMS, which were developed long time age before software integration become important or possible, we created a set of user functions in R that allow executing both models, collecting their outputs and an creating  parameter files for both systems automatically.

With R being able to operate both the GTAP and GTM models automatically, we are able to formulate the combined solution of the both models as a fairly simple math program that is solved in R:

\begin{equation}
\mbox{GTAP}:V_1\to V_2
\end{equation}

\begin{equation}
\mbox{GTM}:V_2\to V_1
\end{equation}

\begin{equation}
s.t. V_2 = V_1
\end{equation}

where $V_1$ is a vector of all variables in both systems, i.e., all variables in GTAP and all variables in GTM, that is solved by the GTAP model (represented by function \mbox{GTAP}) into a vector of variables $V_2$. At the same time, when $V_2$ in processed by the GTM, its solution is $V_1$. To solve the system simultaneously, we require that $V_1$ and $V_2$ be equal. In our case, the vector of variables $V_1$ includes all GTAP variables and the land change variables, which are exogenous to the GTAP model; the output vector of variables $V_2$ contains the change in the rents, which is treated as exogenous in the GTM model. With several iterations, the system converges to the land change and land rent change that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
